---
title: "Homework 3"
author: Chariz Seijo
output: github_document
---

Solution to HW 3. 

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
```

## Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Observations are the levels of items in orders by user. There are user/order variables, such as user ID, order ID, order day, and order hour. There are also item variables, including name, aisle, department, and some numeric codes.

How many aisles, and which are most items from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Make a plot!

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Make a table!

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>%
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

Apples vs. ice cream!

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  knitr::kable()
```


## Problem 2

Load, tidy, and wrangle the data.

```{r}
accel_df = 
  read_csv(
      "./data/accel_data.csv") %>%
  pivot_longer(
    activity.1:activity.1440,
    names_to = "minute",
    names_prefix = "activity.",
    values_to = "activity_counts"
  ) %>% 
  mutate(
    day = factor(day),
    minute = as.numeric(minute))
```


Table (group_by and summarize, explain, factor so day of the week is in correct order)

Plot (ggplot, x = minutes, y = activity count, use geom line, color to indicate day of the week, develop conclusions)


## Problem 3

```{r}
data("ny_noaa")
```

Data cleaning (count, etc.)

2-panel plot (org data first - group by station, year, month + summarize; filter january and july; then plot)

2-panel plot (make each plot, then figure out how to merge; first plot - contour plot or bin plot; second plot - filter first, box plot or ridge plot)